Can We Crunch the Social Graph on the Cheap?

### 2011-01-24 doc/hotpar2011/softxmt-hotpar2011.tex
reviewed evaluation draft in paper

%% move stuff to defs file
\newcommand{\todo}[1]{{\bfseries [[#1]]}}
%% To disable, just uncomment this line
%% \renewcommand{\todo}[1]{\relax}

@@

\section{Current Implementation and Results}
\input{evaluation}

@@

The lists were sized to exceed the last level cache and were laid out
randomly with pointers spaced at a cache line granularity, maximizing
the probability of each reference being a miss. We allocated the lists
on 1GB huge pages to minimize TLB refill overhead. The lists were
allocated in the memory attached to the processor containing the cores
doing the traversal. We ran 10 traversals and recorded the maximum
rate.

@@

Figure~\ref{fig:listwalk-totalconc} shows the result. Each point represents
the maximum rate pointers are traversed for a given number of
concurrent offered references. We see a maximum rate of 277 million
references per second; this is achieved when the number of offered
references is 42. Note that a single core cannot support this level of
memory concurrency; the maximum reference rate for a single core is
107M. We believe this limit is due to the core's {\em line fill
  buffer} \todo{cite nehalem perf docs}, which limits the core to 10
concurrent L2 misses.

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{figures/qpi_bw-edited.pdf}
  \end{center}
  \caption{Pointer chasing in a remote processor's memory.}
  \label{fig:listwalk-qpi}
\end{figure}

Finally, we measured the rate at which the cores can make requests
over the QPI link. For this experiment, we allocated the lists in the
second processor, and ran the traversals on the first
processor. Figure~\ref{listwalk-qpi} shows the result. We found that use
of the QPI link limits us to 175 million references per second.

@@

We observe a gradual decrease in reference rate once the number of
concurrent references per core exceeds 10; we believe this is due to
later prefetches squashing earlier ones in the line fill buffer. We
also observe many samples below the maximum rate even with enough
memory concurrency; this is harder to explain, but we suspect it is
due to resource contention in the cores' pipelines once the memory
system is full of requests.

@@

We ran this code on six cores with all data allocated in the same
processor's memory. We varied the size of the local working set from 2KB
to 4MB, and ran both 0 and 32 local updates for each pointer
traversal. We measured the overall reference rate, including both the

Figure~\ref{fig:cache-pressure} shows the result. On the left of the
figure, we see the list-traversal-only case, where we allocate
per-coroutine space, but make no references to it. As we would expect,
the reference rates grow with the number of coroutines just as in
Figure~\ref{fig:multi-green}. On the right of the figure, we see the
case with 32 updates for each list traversal. While performance
decreases for large working set sizes and large number of coroutines,
we see speedup for up to 128KB working set size with 8 coroutines per
core.

@@

We ran these experiments with two sources of memory concurrency: each
core ran up to 250 coroutines, and each coroutine made between 1 and 4
concurrent references. The results are shown in
Figure~\ref{fig:network-delay}. We are still able to reach a
near-maximum rate of 273 million references per second, but at least 3
concurrent references per coroutine are required. Since a single
reference per coroutine achieves 236 million references per
second. Any network interface would be connected through a QPI link,
which we measured to have a bandwidth limit of 175 million references
per second, our runtime appears to have the capacity to chase pointers
in a remote memory.

\todo{may want to rerun with memory on remote core to incorporate QPI limit?}

>>>>

%% move stuff to defs file
\newcommand{\todo}[1]{{\bfseries [[#1]]}}
%% To disable, just uncomment this line
 %%\renewcommand{\todo}[1]{\relax}

@@

@@

The lists were sized to exceed the last level cache and were laid out
randomly with pointers spaced at a cache line granularity, maximizing
the probability of each reference being a miss. We allocated the lists
on 1GB huge pages to minimize TLB refill overhead. The lists were
allocated in the memory attached to the processor containing the cores
doing the traversal. We ran 10 traversals, measuring average rate in each, and recorded the maximum rate.

@@

Figure~\ref{fig:listwalk-totalconc} shows the result. Each point represents
the maximum rate pointers are traversed for a given number of
concurrent offered references. We see a maximum rate of 277 million
references per second; this is achieved when the number of offered
references is 42 \todo{yes, max is achieved here but essentially levels out earlier like 36; do we want a percentage metric for defining a plateau}. Note that a single core cannot support this level of
memory concurrency; the maximum reference rate for a single core is
107M. We believe this limit is due to the core's {\em line fill
  buffer} \todo{cite nehalem perf docs}, which limits the core to 10
concurrent L2 misses.

@@

% take this figure out if we need the space
% the main point is room for remote references to local memory
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{figures/qpi_bw-edited.pdf}
  \end{center}
  \caption{Pointer chasing in a remote processor's memory.}
  \label{fig:listwalk-qpi}
\end{figure}

Finally, we measured the rate at which the cores can make requests
over the QPI link. For this experiment, we allocated the lists in the
second processor, and ran the traversals on the first
processor. Figure~\ref{fig:listwalk-qpi} shows the result. We found that use
of the QPI link limits us to 175 million references per second.

@@

We observe a gradual decrease in reference rate once the number of
concurrent references per core exceeds 10; we believe this is due to
later prefetches squashing earlier ones in the line fill buffer. We
also observe some repeatable outliers below the maximum rate even with enough
memory concurrency; this is harder to explain, but we suspect it is
due to resource contention in the cores' pipelines once the memory
system is full of requests.

@@

We ran this code on six cores with all data allocated in the same
processor's memory. We varied the size of the local working set from 2KB
to 4MB, and ran both 0 and 32 local updates for each pointer
traversal. We measured the overall reference rate, including both the local updates and remote references.

Figure~\ref{fig:cache-pressure} shows the result. On the left of the
figure, we see the list-traversal-only case, where we allocate
per-coroutine space but make no references to it. As we would expect,
the reference rates grow with the number of coroutines just as in
Figure~\ref{fig:multi-green}. On the right of the figure, we see the
case with 32 updates for each list traversal. While performance
decreases for large working set sizes and large number of coroutines,
we see speedup as number coroutines increases for up to 128KB working set size with 8 coroutines per
core.

@@

We ran these experiments with two sources of memory concurrency: each
core ran up to 167 coroutines, and each coroutine made between 1 and 4
concurrent references. The results are shown in
Figure~\ref{fig:network-delay}. We are still able to reach a
near-maximum rate of 273 million references per second, but at least 3
concurrent references per coroutine are required. Running with a single
reference per coroutine achieves 236 million references per
second. Any network interface would be connected through a QPI link,
which we measured to have a bandwidth limit of 175 million references
per second, so our runtime appears to have the capacity to chase pointers
in a remote memory.

\todo{do we want to include that the memory system imposes max limit so really more throughput is still achievable if there are enough ILP memory concurrency?}
\todo{may want to rerun with memory on remote core to incorporate QPI limit?}
